Random_Effects

### 2023-06-05 Code/BatchEffect.Rmd
Further updates, previous code was wrong

```{r}
nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(nested, one_effect)
batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(batch_year, nested)
#ranef(nested)
```

```{r}
simulation <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z1 <- rnorm(n, sd = vc$sdcor[1])  # random effects for Year
    z2 <- rnorm(n, sd = vc$sdcor[2])  # random effects for Batch
    mu <- exp(fixef(model) + z1 + z2)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
simulation(old_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
```
```{}
All probabilities are for RxnTime=0.1 seconds
old pooled: 0.0283374 vs 0.0057613
men_pooled: 0.0313561 vs 0.0068649
old finals: 0.0056892 vs 0.0004389
men_finals: 0.0083227 vs 0.000856
The times on the left are for the 2 batch, times on the right are for the 1 batch
```

>>>>

```{r}
nested <- glmer(RxnTime ~ (1|Year) + (1|Batch), data=men_pooled, family = Gamma(link="log"))
anova(nested, one_effect)
batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
anova(batch_year, nested)
#ranef(nested)
```

```{r}
old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))

```

```{r}
simulation <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_year <- vc$sdcor[1]
    sd_year_batch <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_year)
    z2 <- rnorm(n, sd = sd_year_batch)

    shape <- 1 / sqrt(sd_year^2 + sd_year_batch^2)
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
simulation(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
simulation(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
simulation(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
simulation(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
```

```{}
All probabilities are for RxnTime=0.1 seconds
old finals: 0.0452759 vs 0.0004389
men_finals: 0.0647663 vs 0.000856
old pooled: 0.0779887 vs 0.0057613
men_pooled: 0.0897533 vs 0.0068649
The times on the left are for the 2 batch, times on the right are for the 1 batch
```
